---
title: "Stat437 Project Analysis - Stroke"
author: "Emily Liu"
date: "3/4/2021"
output: html_document
---

```{r include = FALSE}
library(Matching)
library(tidyverse)
library(tableone)

# read in Prediction of Cerebral Stroke dataset
# Source: https://dataverse.harvard.edu/file.xhtml?fileId=4402679&version=1.0
stroke <- vroom::vroom("~/Projects/Stat437-Project/Strokedataset.csv")
```

# Data cleaning
```{r include = FALSE}

stroke <- na.omit(stroke)

# Change all categorical variables to factors
stroke$gender <- as.factor(stroke$gender)
stroke$heart_disease <- as.factor(stroke$heart_disease)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)

# Create new column "is_self_employed"
stroke$is_self_employed <- stroke$work_type

# Turn is_self_employed variables as binary outcome based off of work_type variables
for(i in 1:length(stroke$is_self_employed)) {
        if(stroke$is_self_employed[i] == "Self-employed") {
                stroke$is_self_employed[i] = 1
        } else {
                stroke$is_self_employed[i] = 0
        }
}

# Change variables in newly created is_self_employed column to factor
stroke$is_self_employed <- as.factor(stroke$is_self_employed)



write.csv(x = stroke, file = "strokeClean.csv")

# Create list of covariates to be used in following Table 1
xvars <- c("gender", "age", "hypertension", "heart_disease", "ever_married", 
           "Residence_type", "avg_glucose_level", "bmi", "smoking_status")
```


# Create Table 1
```{r}
table1 <- CreateTableOne(vars = xvars, strata = "is_self_employed", data = stroke, test = FALSE)
print(table1, smd = TRUE)
```


# Perform logisitc regression
# Examine the distribution of propensity scores for treated and untreated individuals.
```{r}
# logistic regression model
# stroke is outcome (exclude from logistic regression model)
# is_self employed is treatment
reg_model <- glm(data = stroke, is_self_employed ~ gender + age + hypertension + 
                         heart_disease + ever_married + Residence_type + 
                         avg_glucose_level + bmi + smoking_status, 
                 family = binomial())

# Plot propensity scores
pscores <- reg_model$fitted.values

hist_trt <- hist(pscores[stroke$is_self_employed == 1], plot = FALSE)
hist_ctrl <- hist(pscores[stroke$is_self_employed == 0], plot = FALSE)
hist_ctrl$counts = -hist_ctrl$counts
hmax <- max(hist_trt$counts)
hmin <- min(hist_ctrl$counts)
X = c(hist_trt$breaks, hist_ctrl$breaks)
xmax <- max(X)
xmin <- min(X)
plot(hist_trt,
     ylim = c(hmin, hmax),
     xlim = c(xmin, xmax),
     ylab = "Number of people",
     main = "Propensity Scores by Group",
     xlab = "Propensity Score", 
     axes = F)
lines(hist_ctrl)
axis(1, at = seq(0, 1, by = 0.1))
axis(2, at = seq(10, -0, by = -1)) # ***fix axis
```


# Subset your data to consider only treated individuals with a propensity score less than the maximum propensity score of the untreated and consider only untreated individuals with a propensity score greater than the minimum propensity score of the treated individuals. Report the number of individuals excluded.
```{r}
# treated individuals
trt_pscores <- data.frame("self_employed" = pscores[stroke$is_self_employed == 1]) %>% 
        filter(!is.na(self_employed))
min_trt_pscore <- min(trt_pscores) # min for treated

# not treated individuals
ntrt_pscores <- data.frame("not_self_employed" = pscores[stroke$is_self_employed == 0]) %>%
        filter(!is.na(not_self_employed))
max_ntrt_pscore <- max(ntrt_pscores) # max for untreated


# treated individuals < maximum propensity score of the untreated
sum(trt_pscores < max_ntrt_pscore) # none are excluded

# untreated individuals > minimum propensity score of the treated individuals
sum(ntrt_pscores > min_trt_pscore)
nrow(ntrt_pscores) - sum(ntrt_pscores > min_trt_pscore) # 226 were excluded
nrow(trt_pscores) - sum(trt_pscores < max_ntrt_pscore) # 1 was excluded 

```


# Apply propensity score matching and create Table 1
```{r}
# SEE HW 6 FOR REFERENCE

stroke <- read.csv("strokeClean.csv")

set.seed(12345)

logit <- function(p) {
        log(p) - log(1-p)
}

psmatch <- Match(Tr = stroke$is_self_employed, M = 1, X = logit(pscores), replace = FALSE)
matching_df <- stroke[c(psmatch$index.treated, psmatch$index.control),]

matching_t1 <- CreateTableOne(vars = xvars, strata = "is_self_employed",
                              data = matching_df, test = FALSE)
print(matching_t1, smd = TRUE)
```


# Use the matched data set to carry out the outcome analysis.
```{r}

outcome_analysis <- function (df) {
        
trt <- df %>% filter(is_self_employed == 1)
ctrl <- df %>% filter(is_self_employed == 0)

return(t.test(y = trt$stroke, x = ctrl$stroke, paired = TRUE))
        
}
outcome_analysis(matching_df)
```
Self-employed does not make a difference in stroke


# Perform a sensitivity analysis by, for example, adjusting the caliper, the number of matches, or the covariates included in calculating the propensity scores.
```{r}
sens <- Match(Tr = stroke$is_self_employed, M = 1, X = logit(pscores), 
              caliper = 0.0001, replace = FALSE)
sens_df <- stroke[c(sens$index.treated, sens$index.control),]
sens_t1 <- CreateTableOne(vars = xvars, strata = "is_self_employed",
                              data = sens_df, test = FALSE)
outcome_analysis(sens_df)
print(sens_t1)
```





